---
title: "Data Science on Education with .red[**R**]"
subtitle: "First Class"
author: "Yangyong Ye"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:9
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false

---


# 为什么学数据科学与教育？

- ###    数据无处不在(Data, Information, Knowledge, and Wisdom)

- ###    研究走向实证(describe、explain、predict、intervene)

- ###    日常工作需要


???
（各种工作都会面临着数据处理与信息加工问题），数据处理已经成为一种基本生存技能


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
```

---
class: center, middle,inverse
> Wisdom is not a product of schooling but of the lifelong attempt to acquire it.
> -----Albert Einstein


---
class: center, middle,inverse

# 讨论
## 请举一个急待解决的教育问题


???
- 缺少编制？

- 工资待遇低？

- 高级职称指标少？

- 工作任务繁重？

- 新教师质量越来越差？
可以请学生罗列他们认为教育领域存在哪些棘手的问题？
2021.11.12,note:这个练习不好，学生提出的问题，无法引入到说明数据分析比较重要这一主题上来。因此，以后要在此环节
设计几个因没有注重数据分析而带着教育决策失误的例子。

---

# 数据科学研究什么

## - 数据收集
## - 数据存储与管理
## - .red[数据分析]
## - *数据呈现*

---
# 数据科学与统计学

## - 数据科学以统计学为基础

## - 数据科学应用性更强（机器学习、因果推断）

---
# 定量研究步骤

- 确定主题
   - 目标群体（测量单位）
   - 相关变量
- 理论构架（提出假设）
- 收集数据
  - 问卷调查
  - 二手数据（官方数据、共享数据）
- *数据清理*
- 模型建构
  - 模型选择
  - 变量选择
- *数据分析与展示*
- 结果解读
- 启示建议


---
class: inverse, middle

# Course Overview

## - Environment Preparation

## - Data fetching

## - Data preparation

## - Data tabulation

## - Data visualization

## - Data analysis


---
class: inverse, middle

# Things not Covored in the Class

## - statistics
## - data collection
## - machine learning
## - causal inference

---
class: center middle

# Class Schedules

```{r,echo=FALSE, fig.width=10}
library(timevis)
classplan <- data.frame(
  id = 1:8,
  content =c("week 1: \nintroduction", "week 2: data preparation", "week 3: data tabulation", "week 4: data visualization I",
             "week 5: data visualization II", "week 6: data analysis", "week 7: data analysis", "week 8: data analysis"),
  start= c("2020-11-9","2020-11-16","2020-11-23","2020-11-30","2020-12-7",
           "2020-12-14","2020-12-21","2020-12-28")
)


timevis(classplan,
        options = list(width = "800px",height="400px"))
```

---
class: center, middle,inverse
# 如何学好这门课？

## - 不要恐惧

## - 加强自学

## - **练习、练习、练习**

---
class: inverse, middle

# Environment Preparation

##  - RStudio installation
  
##  - RStudio introduction
  
##  - R introduction
  
##  - Common package installation (eg. **tidyverse**)
  
##  - Self-learning material [R for data science](http://r4ds.had.co.nz/)

---
class: inverse, middle

# Data Fetching

##  - 中国教育追踪调查[CEPS](http://ceps.ruc.edu.cn/)
  
##  - 中国家庭追踪调查[CFPS](http://www.isss.pku.edu.cn/cfps/)
  
##  - 中国社会综合调查[CGSS](http://cgss.ruc.edu.cn/)
  
##  - 中国家庭收入调查[CHIP](http://ciid.bnu.edu.cn/chip/index.asp)
  
##  - 中国健康与营养调查[CHNS](https://www.cpc.unc.edu/projects/china)

---
class: inverse, middle

# 熟悉调查数据的步骤

## - user guide

## - questionnaire

## - technical report

## - codebook

## - data


---
class: inverse, middle
# 官方教育统计数据

## - 中国教育年鉴（1949-1981，1982-1984，1985-1986,1988）[EDU](http://cyfd.cnki.com.cn/N2014080445.htm)
  
## - 中国教育统计年鉴(1987-2017)[EDU_STAT](http://data.cnki.net/yearbook/Single/N2017120244)
  
## - 中国教育经费统计年鉴[EDU-FINA](http://data.cnki.net/yearbook/Single/N2018070117)

---
class: inverse, middle
# 练习
为了让大家熟悉教育统计数据，现需要大家查找以下数据
1.分省教育经费支出明细（小学）
2.分省教育经费支出明细（初级中学）
3.分省小学专任教师数（小学）
4.分省中学专任教师数（初级中学）
分工：
张韵歌（2017－2019）
张菁（2014－2016）
张晓帆（2011－2013）
田达廷（2008－2010）
千艾灵（2005－2007）
黄洋子（2002－2004）
王思祯（1999－2001）
提示：1.支出明细数据可在《教育经费统计》；2.专任教师数可在国家统计局官网或者其它途径。3.数据上课前交给黄洋子。


---
class: middle,inverse
# Data Preparation

##  - recode/rename
  
##  - subset(select/filter)
  
##  - join/combine
  
##  - arrange
  
##  - summarize
  
##  - count
  
##  - reshape

---
class: middle,inverse

# Data Tabulation

###  - frequency table
  
###  - cross-table
  
###  - summary table
  
###  - test-table(t, chi, anovo)
  
###  - regression table


---
class: inverse, middle

# Data Visualiztion[Graph Type](https://www.r-graph-gallery.com/)

###  - **Amount**: Barplot,circular barplot, Spider/Radar, parallel, lollipop/stem, circular barplot
  
###  - **Distribution**: histgram, density, boxplot, violin, ridgeline/joyplot
  
###  - **Correlation**: scatter, heatmap, correlogram, bubble, connected scatter, density 2D
  
###  - **Composition**: treemap, dendrogram, venn diagram, stacked bar, pie chart, doughnut, circular packing

---
class: inverse, middle

# Data Visualiztion[Graph Type](https://www.r-graph-gallery.com/)

###  - **Trending**: line, Area, stacked area, streamgraph, parallel, time series
  
###  - **Maps**: background map, connection, choropleth, bubble, hexbin map, cartogram
  
###  - **Flow**: network, chord diagram, hierarchical edge bundling, sankey diagram
  
###  - **Other**: interactive, animation, 3d, color, cheat sheet.

---
class: middle,inverse

# Data Visualiztion - Useful packages

###  - comprehensive package: **ggplot2**

###  - for dendrogram: **packcircles**,**ggraph**, **treemap**, **dendextend** 

###  - for network: **igraph**,**tidygraph**,**ggraph**

###  - correlation matrix: **ellipse**, **car**, **GGally**,**corrgram**

###  - flow diagram: **ggalluvial**


---
class: inverse, middle,center

# Intentionally Blank

---
class: middle,inverse
# Data Analysis

###  - simple test(**t**, *chi*, anova)

###  - correlation
  
###  - factor analysis
  
###  - cluster analysis

###  - regression

---
class: middle,inverse

# Logistics - MATERIALS

###  - Garrett Grolemund, Hadley Wickhan. [R for data science](https://r4ds.had.co.nz/)
###  - [An Introduction to R](https://intro2r.com/)
###  - [Introduction to Data Science](https://rafalab.github.io/dsbook/)
###  - [ggplot2-Elegant-Graphics for Data Analysis(2nd)](https://ggplot2-book.org/)
###  - 王孝玲,教育统计学,华东师范大学出版社.
###  - [open access materials](https://bookdown.org/)

---
class: middle,inverse

# Logistics- ACCESSMENT
###  - Presence and class performance      (10%)
###  - Assignments                         (40%)
###  - Individual project proposal         (50%)
 
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```
class: center, middle,inverse

# Let's begin ou.red[**R**] journey!

---
class: center, middle,inverse

# Why Language?

## Fast
## Powerful
## Note-taking
## Reproducible

---
class: center, middle,inverse

# Why .red[**R**]?
## small
## versatile
## beautiful
## not hard to learn

???

In this slide, I will talk about the advantages of R and why I change from spss to R.
  
---
class: middle,inverse, center

# Project Example

## 教育经费情况分省报告
## 不同年代的人受教育程度
## 高考成绩分析报告
## 技能需求结构变化
## Teacher self-efficacy

???
高考成绩分析三个项目：（1）语数外成绩分布的问题：为什么语文的变异小数学的变异大？为什么英语不呈现正太分布？为什么数学左边是锯齿状的？哪门学科更重要？
（2）不同群体的对比问题，男女、城乡、应往届、文理科
（3）入学年龄对成绩的影响

---

# Introduction to R

### - R is a language and environment for statistical computing and graphics.

### - R is available as Free Software.

### - R can be extended (easily) via packages.


---
# Example in R window

`r chunk_reveal("plotex")`


```{r, plotex}
data(cars)
head(cars)
with(cars, plot(speed, dist))
```

---

# Intro to Rstudio

- Intro panel: source/console/environment

### 1. create a script, and save it, and open
 
### 2. create a project, put a data in it
 
### 3. install package, like “tidyverse”, it is important to change mirror source
 
### 4. options/appearance, setup theme.
 
---
# Rstudio package installation

 
`r chunk_reveal("plotex2")`

```{r,plotex2, echo=FALSE}


library(tidyverse)
ggplot(mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

---
class: middle, inverse

# A Swiss-Army Knife for Data I/O [rio](https://cloud.r-project.org/web/packages/rio/vignettes/rio.html)


```{r, plotex3,echo=TRUE}
#install.packages("rio")
library(rio)
install_formats()
```

---

# Homework for .red[NEXT] Week

### - Chapter 1-4 from Doc. Wang

### - Chapter 1-8 from [R4DS](https://r4ds.had.co.nz/) (required)

### - Chapter 1- 3 [Introducation to R](https://intro2r.com/install-r.html?s=09)(recommended)

### - Chapter 10 [Efficient R programming](https://csgillespie.github.io/efficientR/learning.html)

### - Download **CEPS** data. read documentations


---
# Packages need to be installed

### - data import: rio

### - data manipulation: tidyverse; sjPlot, sjmisc, janitor, skimr, naniar, visdat,

