---
title: "Data Preparation"
subtitle: "second class"
author: "Yangyong Ye"
institute: "SOE, RUE"
date: "2020/11/16 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [chocolate, chocolate-fonts,my_css.css]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: 16:9
---


```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F)
library(flipbookr)
library(tidyverse)
library(rio)
student <- import("/Users/yangyongye/microcloud/projects/survey_data/ceps/2014-2015data/cepsw2studentCN.dta")
```


# Rstudio Intro
  
- Panel: source/console/environment/files/plots/packages/help/viewer

- Exercise: 

  1. create a script, save, close and open it

  2. create a project, put a data in it.

  3. install package, like "tidyverse", load it.

  4. options/appearance, setup theme, it is important to change **mirror source**.

---
# Data I/O

+ readr: reads .csv data.

+ haven: reads SPSS, Stata, and SAS files.

+ readxl: reads excel files (both .xls and .xlsx).

+ read_json: reads json data.

+ read_ogr: write_rds() and read_rds() are uniform wrappers around the base functions readRDS() and saveRDS().

+ **rio**: .red[import, export, convert]

---

# Commonly Used Functions

- ?
- str
- summary
- %in%
- head, tail
- with
- round, n, sum, mean, sd
- length, dim, ncol, nrow, names, colnames, rownames
- rep, seq, rev, sample, ifelse, list

---

# [tidyverse](https://www.tidyverse.org) packages:

- ggplot2, for data visualisation.

- dplyr, for data manipulation.

- tidyr, for data tidying.

- readr, for data import.

- purrr, for functional programming.

- tibble, for tibbles, a modern re-imagining of data frames.

- stringr, for strings.

- forcats, for factors.

---
# Colummn Manipulation Overview

  - select
  
  - rename
  
  - relocate
  
  - mutate
  
  - recode
  
---

# Overview of selection features [tidyselect](https://tidyselect.r-lib.org/reference/language.html)

Tidyverse selections implement a dialect of R where operators make it easy to select variables:

- : for selecting a range of consecutive variables.

- ! for taking the complement of a set of variables.

- & and | for selecting the intersection or the union of two sets of variables.

- c() for combining selections.

In addition, you can use selection helpers. Some helpers select specific columns:

- everything(): Matches all variables.

- last_col(): Select last variable, possibly with an offset.

---

# Overview of selection features [tidyselect](https://tidyselect.r-lib.org/reference/language.html)
- These helpers select variables by matching patterns in their names:

  - starts_with(): Starts with a prefix.

  - ends_with(): Ends with a suffix.

  - contains(): Contains a literal string.

  - matches(): Matches a regular expression.

  - num_range(): Matches a numerical range like x01, x02, x03.

- These helpers select variables from a character vector:

  - all_of(): Matches variable names in a character vector. All names must be present, otherwise an out-of-bounds error is thrown.

  - any_of(): Same as all_of(), except that no error is thrown for names that don't exist.

- This helper selects variables with a function:

   - where(): Applies a function to all variables and selects those for which the function returns TRUE.


---

# Row manipulation

  - filter
  
  - arrange
  
  - slice: slice_head, slice_tail, slice_sample, slice_min, slice_max


---
# filter: conditions

- R provides the standard suite: >, >=, <, <=, != (not equal), and == (equal).

- Boolean operators yourself: & is ??and??, | is ??or??, and ! is ??not??.,


---
# R basic:

[Data type](https://evamaerey.github.io/data_manipulation/about)

---

# Multiple column or row manipulation

  - across
  
  - group_by & summarize
  
  - count
  
  - distinct
  
  - rowwise & c_across


???

this is not right
  
---

# Separate & unite

  - extract
  
  - separate
  
  - unite

---

# TWO TABLE VERBS

- Mutating Joins 
  — inner_join(), left_join(), right_join(), full_join()

- Filtering Joins 
  — semi_join(), anti_join()

- Set Operations — union(), intersect(), setdiff()
  - intersect(x, y): return only observations in both x and y
  
  - union(x, y): return unique observations in x and y
  
  - setdiff(x, y): return observations in x, but not in y

---

`r chunk_reveal("join")`

```{r join, include=FALSE}
library(gapminder)
library(tidyverse)
gapminder %>% 
  select(country, lifeExp, year) -> 
  country_life
gapminder %>% 
  select(country, pop, year) -> 
  country_pop
country_pop %>% 
  full_join(country_life,
            by = c("country","year"))
```
---


# join
`r chunk_reveal("lef", widths =c(50,50))`


```{r lef, include=FALSE}
library(nycflights13)

# mutating joins: left_join,right_join, full_join, inner_join
flights %>%
  select(year:day, hour, origin, dest, tailnum, carrier) -> flights2
flights2 %>%
  left_join(airlines, by="carrier")

flights2 %>%
  left_join(weather,
            by=c("origin","year","month","day"))
flights2 %>%
  left_join(airports, c("dest"="faa"))
```


---
# anti & semi
`r chunk_reveal("lef2")`


```{r lef2, include=FALSE}
flights %>%
  count(dest, sort = TRUE) %>%
  head(10) -> top_dest

flights %>% semi_join(top_dest)
flights %>% anti_join(top_dest)

flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(tailnum, sort = TRUE)

```

---

# long & wide [tidyr](https://tidyr.tidyverse.org/articles/pivot.html)


- pivot_longer()

- pivot_wider()

---

`r chunk_reveal("long_wide")`

```{r long_wide, include=FALSE}
gapminder %>%
  select(country, continent, lifeExp, year) %>% 
  pivot_wider(names_from =year,
              values_from = lifeExp) ->
  gapminder_life_exp_wide
gapminder_life_exp_wide %>% 
  pivot_longer(`1952`:`2007`,
               names_to = "year",
               values_to = "lifeExp") %>% 
  mutate(year=as.numeric(year)) ->
  gapminder_life_exp_long
```

---
# Practice

* 学生id, 学校id， 城市id，是否独生子女、是否近视、是否恋爱、期中考试语文原始成绩、是否实行百分制

* 只选择期中语文成绩实行100制的学生

* 将语文成绩的变量名称更改为chinese, 将是否独生子女变量名称更改为child_only, 是否恋爱变更为in_love.

---

# First Impression

- sjPlot:: view_df()

- dplyr:: glimpse()

- skimr::skim()

- janitor::clean_names() & remove_empty()

---

# Missing value

- visdat::vis_miss()

- tidyr::drop_na()

- tidyr::replace_na()

- naniar::replace_with_na()

- base::is.na()

---
# Missing value [naniar](http://naniar.njtierney.com/)

Shorthand summaries for missing data:
- n_miss() and n_complete()

- pct_miss()and pct_complete()

Numerical summaries of missing data in variables and cases:
- miss_var_summary() and miss_var_table()

- miss_case_summary(), miss_case_table()


---
class: center, middle

# factor R4DS [15](https://r4ds.had.co.nz/factors.html)

---
class: center, middle
# date and time R4DS [16](https://r4ds.had.co.nz/dates-and-times.html)

---
class: inverse middle
# NEXT WEEK

- [gtsummary](http://www.danieldsjoberg.com/gtsummary/)

- [rstatix](https://rpkgs.datanovia.com/rstatix/)

- [flextable](https://davidgohel.github.io/flextable/index.html)

- [reactable](https://glin.github.io/reactable/)

- install_packages: skimr, janitor, visdat, naniar, rstatix, gtsummary, flextable, DT,


???

教学反思，学生普通对于R语言熟练程度和理解的程度还比较低，细小错误很多，影响了课程的进度，应当在课前明确课下预习的内容。并要求学习做必要练习,这样才能提高课程的进度与成效。

---
class: inverse, middle, center

# Thanks !
